---
title: "Assignment 1"
author: "Maia Cossette"
date: "2025-09-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part 1
#1 Creating a species vector and randomly sample it 100 times
```{r}

set.seed(1)
species<- c("cat", "dog", "whale", "dolphin", "shark")
sample(species, size = 100, replace = TRUE)
```

#2 Creating a width vector and using a log distribution to avoid negatives or zeros
```{r}
set.seed(1)
Limb_width<- c(1:100)
rlnorm(100, meanlog = 0, sdlog =1)

```

#3 Created a length vector and used a normal distribution 
```{r}
set.seed(1)
Limb_length<- rnorm(100, mean = 50, sd = 10)
Limb_length
```

#4 Created a names vector with 100 sample
```{r}
set.seed(1)
Observer<- c("Maia", "Peter", "Jade")
Observers<- sample(Observer, size = 100, replace = TRUE)
Observers

```

#5 Created a data frame with all the vectors and turned it into csv file
```{r}
measurements<-data.frame(species, Limb_width, Limb_length, Observers)
write.csv(measurements,file= "measurements", row.names = F)
head(measurements)
```

#Part 2
```{r}
#upload library 
library(dplyr)

#upload data from r script 1
Data1<- read.csv("measurements.csv")
Data1

#add volume column 
Data1<- Data1 %>%
  mutate(Volume= Limb_width*Limb_length)
Data1

#overwrite new data
write.csv(Data1, "meansurements.csv", row.names = F)
```

#sort data by species, observer and volume
```{r}

Data1<-Data1 %>%
  arrange(species, Observers, Volume)
head(Data1)
```

#average volume for each species 
```{r}
average<-Data1 %>%
  group_by(species)
average
```

#table for number of observations
```{r}
nobservations<- Data1 %>%
  group_by(species, Observers)
nobservations
```

#boxplot 
```{r}
library(ggplot2)
ggplot(data = Data1, aes(x= species, y= Volume)) +geom_boxplot() +labs(title = "Distribution of Limb Volume by Species", x= "Species", y= "Limb Volume")

```
#multipanel plot

```{r}
ggplot(data = Data1, aes(x= Volume)) +geom_histogram() +labs(title = "Distribution of Limb Volume per Species", x= "Volume", y= "Frequency") +facet_wrap(~species)

#facet_wrap for seperate histogram plots 

```

